LWE Trapdoor One-Way Function

### 2017-06-26 latex/qtpie/paper.tex
added key agreement idea
random walks are cool :)

\author{Ella Rose}
\institute{Paso Robles, CA \email{python_pride@protonmail.com}}
\title[EPQ Trapdoor Function]{EPQ Trapdoor Function}

@@

\begin{abstract}
 We define a simple trapdoor function that is suitable for use as a key exchange algorithm. The algorithm is relatively efficient in terms of key size, key generation time, ciphertext size, and public/private key operation time: The public key operation utilizes a single multiplication and addition modulo a big prime.
 \end{abstract}

@@

\section{Introduction}
A Trapdoor One-Way Function (TOWF) is a function that is easy to compute in one direction, but is hard to compute in reverse without some additional piece of information. A TOWF makes exchanging cryptographic keys a straightforward exercise: Alice can simply apply a TOWF to a randomly generated key and send the output to Bob, who knows the trapdoor information. Bob can then use the trapdoor information to invert the output and obtain the key.

In this paper we define a candidate TOWF that is suitable for use to exchange cryptographic keys. It is based on a simple observation of a congruence, $pq + e \equiv q + p_i e \mod N$, along with some subtle manipulation of the sizes of the numbers involved.

\section{Definitions}
We use $+$ to denote addition, $a / b$ to denote division, and $\mod p$ to denote the modulo operation. We use both $a b$ and $a * b$ to indicate multiplication. We use the function $inverse(k, n)$ to indicate the computation of the modular multiplicative inverse of $k \mod n$. We use the function $random(k)$ to indicate the generation of of a random $k$-bit integer.

\section{Algorithm}
There is a publicly agreed upon modulus $N$. The same $N$ may always be used. It must be prime; It it is not secret, and otherwise only needs to be both agreed upon and large enough. The size of $N$ is determined by the rest of the parameter sizes which are discussed below; In our python implementation which targets a 256-bit security level, $N$ is slightly less then 800 bits in size. The larger $N$ is, the longer the BigNum arithmetic operations take, and the larger ciphertexts become; Consequently, we opt for the smallest $N$ possible.

@@

\begin{flalign*}
p_i := inverse(p, N)\\
t := p_i c \mod N\\
q := t \mod p_i\\
t := t - q \\
e := t / p_i
\end{flalign*}

@@

@@

\section{Security Analysis}
There are two main security goals: First, given the public key, the private key must not be obtainable; Second, The output of the public key operation must not be invertible without the private key.

@@

\subsection{Public Key Operation}
The security of the public key operation appears to scale with the size of the randomly generated $q$. It remains to be proven that given $c := p q + e \mod N$, $p$, recovering $e$ is difficult.

>>>>

\author{Ella Rose}
\institute{Paso Robles, CA \email{python_pride@protonmail.com}}
\title[LWE Trapdoor One-Way Function]{LWE Trapdoor One-Way Function}

@@

\begin{abstract}
 We define a simple trapdoor function based on the Learning-With-Errors problem in one dimension, and present a key exchange algorithm based on the trapdoor. The trapdoor algorithm is efficient in terms of key size, key generation time, ciphertext size, and public/private key operation time: We can perform many thousands of public and private key operations per second with a naive python implementation on modest hardware. 
 \end{abstract}

@@

\section{Introduction}
A Trapdoor One-Way Function (TOWF) is a function that is easy for anyone to compute in one direction, but is hard to compute in reverse without some additional piece of information. A TOWF makes exchanging cryptographic keys a straightforward exercise: Suppose that Bob happens to know the trapdoor information for a TOWF: Alice can simply apply the TOWF to a randomly generated key and send the output to Bob. Bob can then use the trapdoor information to invert the output and obtain the key.

The Learning-With-Errors (LWE) problem is a mathematical problem that is believed to be difficult to solve efficiently even with a quantum computer.

In this paper we define a candidate TOWF based on 1-dimensional LWE. It is based on a simple observation of a congruence, $pq + e \equiv q + p_i e \mod N$, along with some subtle manipulation of the sizes of the numbers involved.

\section{Definitions}
We $+$ to denote addition, $a b$ to indicate multiplication, $a / b$ to denote division, and $\mod p$ to denote the modulo operation. We use the function $inverse(k, n)$ to indicate the computation of the modular multiplicative inverse of $k \mod n$. We use the function $random(k)$ to indicate the generation of of a random $k$-bit integer.

\section{Algorithm}
There is a publicly agreed upon modulus $N$. The same $N$ may always be used. It should be prime; It it is not secret, and otherwise only needs to be both agreed upon and large enough. The size of $N$ is determined by the rest of the parameter sizes which are discussed below; In our python implementation which targets a 256-bit security level, $N$ is slightly less then 800 bits in size. The larger $N$ is, the longer the BigNum arithmetic operations take, and the larger ciphertexts become; Consequently, we opt for the smallest $N$ possible.

@@

\begin{flalign*}
p_i := inverse(p, N)\\
t := p_i c \equiv p_i p q + p_i e \equiv q + p_i e \mod N\\
q := t \mod p_i\\
t := t - q \\
e := t / p_i
\end{flalign*}

@@

\subsubsection{Processing the Secret}
The secret should be processed before use a key. Ideally, HKDF should be used; If it is not available, a simple hash function applied to the secret should be sufficient.

@@

\section{Security Analysis}
There are two main goals: First, given the public key, the private key must not be obtainable; Second, The output of the public key operation must not be invertible without the private key.

@@

\subsection{Public Key Operation}
The public key operation appears to be a 1-dimensional version of the learning-with-errors problem.\\

If we consider the case where $pr = 1$, then $prq + e \equiv q + e \mod P$ for uniformly random values $q, e$.
